Kindly Bent to Free Us

### 2019-06-28 kindly/motivation.tex
Add the 4th "big" code example.

\subsection{Type abstraction and linearity violations}

While linearity allows to verify many properties, it also prevents
the implementation of functions that are nevertheless safe. Rust addresses this
issue through the \lstinline/unsafe/ construct, which allows to locally
ignore linearity constraints. This is essential to build basic
functions such as splitting strings.
%
\lang allows to partially emulate this behavior using type abstraction
as provided by ML modules.
In particular, data types can be internally defined as unrestricted but exposed
as affine.

\TODO{Show example with arrays}

@@

\TODO{Show a queue attached to a file}

>>>>

% \subsection{Type abstraction and linearity violations}

% While linearity allows to verify many properties, it also prevents
% the implementation of functions that are nevertheless safe. Rust addresses this
% issue through the \lstinline/unsafe/ construct, which allows to locally
% ignore linearity constraints. This is essential to build basic
% functions such as splitting strings.
% %
% \lang allows to partially emulate this behavior using type abstraction
% as provided by ML modules.
% In particular, data types can be internally defined as unrestricted but exposed
% as affine.

% \TODO{Show example with arrays}

@@

We have seen that borrows are limited to a region. In the various examples so
far, such regions didn't appear in the type signature explicitly.
We know present an example implement of a \texttt{Pool} module
that uses such region to enforce linearity of the objects in the pool.

\TODO{finish explanation}

\begin{figure}[!h]
  \centering
  \begin{subfigure}[t]{.45\linewidth}
\begin{lstlisting}
type ('a: 'k) pool : 'k = {
  spawn : unit -> 'a ;
  queue : 'a queue;
}

let create spawn =
  { spawn ; queue = Queue.create () }
let consume f c = Queue.iter f c.queue

let use { spawn ; queue } f =
  let o = match Queue.pop &queue with
    | Some x -> x
    | None () -> spawn ()
  in
  let r = f &!o in
  Queue.push o &queue;
  r
\end{lstlisting}
    \caption{Implementation}
  \end{subfigure}~
  \begin{subfigure}[t]{.55\linewidth}
\begin{lstlisting}
type ('a: 'k) pool : 'k

create : (unit -> 'a) -> 'a pool
consume : ('a -> unit) -> 'a pool -> unit
use : ('a_1:'k_1),('a_2:'k_2),('k_1 < lin_r),('k_2 < lin_r) =>
  &('a_1 pool) -> (&!(aff_r,'a_1) -{'k}> 'a_2) -{'k_1}> 'a_2
\end{lstlisting}
    \caption{Interface}

\begin{lstlisting}
let uri = "postgresql://localhost:5432"

(* Create a database connection pool. *)
let pool = Pool.create (fun () -> Db.connect uri)

(* Use the pool in queries. *)
let create_user name =
  Pool.use &pool (fun connection ->
    Db.insert "users" [("name", name)] connection)
\end{lstlisting}
    \caption{Example of use}
  \end{subfigure}

  \caption{The \texttt{Pool} module}
\end{figure}
